MLH1 Results Outline

### 2020-02-07 doc/Results.rmd
added some results for Nrm Foci Pos

```{r male.Nrm1CO_strain, echo=FALSE, eval=FALSE}

#isolate all musc
Curated_BivData_male <- Curated_BivData[Curated_BivData$sex == "male",]

#make table
Musc_male_1COpos_table <- ddply(Curated_BivData_male[Curated_BivData_male$hand.foci.count == 1,], c("strain"), summarise,

                        n_obs = length(Obj.ID),
                         mean_F1 = mean(Foci1, na.rm=TRUE),
                        mean_F1.nrm = mean(PER_Foci_1,na.rm=TRUE)

@@

#Dom
glm_SC.male_Dom <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])

#summary(glm_SC.male_Dom)  #ALL strain are sig
#both G and LEW have significant mean.SC, doesn't meet predictions

#Musc
glm_SC.male_Musc <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])

#summary(glm_SC.male_Musc)

#Mol
glm_SC.male_Mol <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])

#summary(glm_SC.male_Mol)

@@

@@

log.SC_male_3CO.Mol <- glm(Rec.group ~ mean.SC_3CO,
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
#summary(log.SC_male_3CO.Mol)

@@

@@

## Normalized CO positions

>>>>

```{r male.Nrm1CO_strain, echo=FALSE, eval=FALSE}
#isolate all musc
Curated_BivData_male <- Curated_BivData[Curated_BivData$sex == "male",]

#make table
Musc_male_1COpos_table <- ddply(Curated_BivData_male[Curated_BivData_male$hand.foci.count == 1,], c("strain"), summarise,
                        n_obs = length(Obj.ID),
                         mean_F1 = mean(Foci1, na.rm=TRUE),
                        mean_F1.nrm = mean(PER_Foci_1,na.rm=TRUE)
                         )

##long biv
male.long.biv <- real.long.bivData %>% filter(sex== "male") %>% filter(species == "M.musculus") 
#table( droplevels( male.long.biv$strain)  )

mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,

                        n_obs = length(Obj.ID),
                         mean_F1 = mean(Foci1, na.rm=TRUE),
                        mean_F1.nrm = mean(PER_Foci_1,na.rm=TRUE),

                        mean_SC = mean(chromosomeLength),

                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE)

                                               )
#add subsp

mouse.av.long.biv.table <- add_strain(mouse.av.long.biv.table)
mouse.av.long.biv.table <- add_subsp(mouse.av.long.biv.table)

mouse.av.long.biv.table$strain <- factor(mouse.av.long.biv.table$strain, ordered = FALSE)
mouse.av.long.biv.table$subsp <- factor(mouse.av.long.biv.table$subsp, ordered = FALSE)

#add rec group
mouse.av.long.biv.table$Rec.group <- ifelse(grepl("PWD", mouse.av.long.biv.table$strain), 1, 
                            ifelse(grepl("MSM", mouse.av.long.biv.table$strain), 1,
                             ifelse(grepl("SKIVE", mouse.av.long.biv.table$strain), 1, 0)) )

@@

#Dom
glm_SC.male_Dom <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
#summary(glm_SC.male_Dom)  #ALL strain are sig
#both G and LEW have significant mean.SC, doesn't meet predictions

#Musc
glm_SC.male_Musc <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
#summary(glm_SC.male_Musc)

#Mol
glm_SC.male_Mol <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
#summary(glm_SC.male_Mol)

@@

```{r Q2.SC.long.biv, echo=FALSE}

#mouse.av.long.biv.table

glm1.SC.male_long.biv <- glm(mean_SC ~ subsp * strain, data= mouse.av.long.biv.table)
summary(glm1.SC.male_long.biv)

glm2.SC.male_long.biv <- glm(mean_SC ~ strain, data= mouse.av.long.biv.table)
summary(glm2.SC.male_long.biv)

#Dom
glm_SC.male_Dom_long.biv <- glm(mean_SC ~ strain, data= mouse.av.long.biv.table[mouse.av.long.biv.table$subsp == "Dom",])
summary(glm_SC.male_Dom_long.biv)  #ALL strain are sig
#both G and LEW have significant mean.SC, doesn't meet predictions

#Musc
glm_SC.male_Musc_long.biv <- glm(mean_SC ~ strain, data= mouse.av.long.biv.table[mouse.av.long.biv.table$subsp == "Musc",])
summary(glm_SC.male_Musc_long.biv)

#Mol
glm_SC.male_Mol_long.biv <- glm(mean_SC ~ strain, data= mouse.av.long.biv.table[mouse.av.long.biv.table$subsp == "Mol",])
summary(glm_SC.male_Mol_long.biv)

## logistic regression for Rec group

log.SC.male_long.SC <- glm(Rec.group ~ mean_SC, 
              data=mouse.av.long.biv.table,  family=binomial(link="logit"))

summary(log.SC.male_long.SC)

@@

#there's no rec group for this table
log.SC_male_3CO.Mol <- glm(Rec.group ~ mean.SC_3CO,
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
#summary(log.SC_male_3CO.Mol)

#reverse logistic regression (single bivalent level)
#how well does SC length predict 1 or 2 COs?

Bivalent_1o2 <-  Curated_BivData_male %>% filter(species== "M.musculus") %>% filter(hand.foci.count > 0) %>% filter(hand.foci.count < 3)

#need to switch the co counts to 1 - 0
Bivalent_1o2$chrm.class <- ifelse(grepl(1, Bivalent_1o2$hand.foci.count), 0, 
                                    ifelse(grepl(2, Bivalent_1o2$hand.foci.count), 1, ""))

Bivalent_1o2$chrm.class <- as.numeric(Bivalent_1o2$chrm.class)
Bivalent_1o2$strain <- factor(Bivalent_1o2$strain, ordered = FALSE)

log.SC_male_chrm.class <- glm(chrm.class ~ chromosomeLength*strain,
              data=Bivalent_1o2,  family=binomial(link="logit"))

summary(log.SC_male_chrm.class)
#not sure if this is the equivilent of running each strain seperatly
#the strain effect is probably also pulling in the different proportion of chrm class
#useing the chrm.length*strain gives fewer significant factors.

#i'd be more convinced by running all those log regression plots for each strain
#come back later
```

```{r Q2_SC_ChrmClass_long.biv, echo=FALSE}
#there might not be enough data 
#
ii <- as.data.frame( table(male.long.biv$hand.foci.count, male.long.biv$strain) )
```

@@

# Q2 Normalized CO positions

<!-- 
C.1) not enough is known about variation within species for the  **1CO normalized positions**. Null prediction, no difference in the 'telomeric pattern'.

C.3) telomere and centromere distances ... 
-->

Since there aren't many good predictions for how the 1CO normalized landscape relates to gwRR variation, these will be the main questions

- Has there been evolution in the 1CO pattern?

- Is there a consistant pattern with the high rec strains?

Normalized 1CO foci positions will be used. (F1_PER). Check fixed effects of subsp and strains.

sis-co-ten, telomere and centromere distance of foci are metrics which draw from a wider pool of samples. (So maybe I will use them... but avoiding focus on them)

```{r Q2.show.nrmF1, echo=FALSE}
#Curated_BivData_male

#boxplots of nrm F1
male.bivdata.1CO <- Curated_BivData_male %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)

Q2.show.nrmF1 <- ggplot(male.bivdata.1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter(aes(alpha=.5) )+ geom_boxplot(aes(y= PER_Foci_1, alpha=.5) ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Normalized 1CO Foci pos")

Q2.show.nrmF1

nrmF1_mouse_level <- ggplot(Male.poly.Mouse.Table_BivData_4MM, aes(x=strain, y=norm.first.foci.pos, color=strain) )+geom_jitter(aes(alpha=.5) )+ geom_boxplot(aes(y= norm.first.foci.pos, alpha=.5) ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse Level, \nNormalized 1CO Foci pos")

nrmF1_mouse_level

Remade the Position table, -- and MSM has a more telomeric distribution. MOLF has the most central distribution.

```{r Q2.F1.pos.long.biv, echo=FALSE}

#make the same plots as above but with the long biv data set
male.long.biv.1CO <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)

#mouse.av.long.biv.table

Q2.show.nrmF1_long.biv <- ggplot(male.long.biv.1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter(aes(alpha=.5) )+ geom_boxplot(aes(y= PER_Foci_1, alpha=.5) ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Normalized 1CO Foci pos \nLong Biv")

Q2.show.nrmF1_long.biv

#2 version of the boxplot data -- not sure why th pattern differd

#more bivData
#nrmF1_with.SC.length <- ggplot(Curated_BivData_male, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter(aes(alpha=.5) )+ geom_boxplot(aes(y= PER_Foci_1, alpha=.5) ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Normalized 1CO Foci pos")

I think the plot shows that there are fewer 1CO for the High Rec group.

Maybe showing showin

```{r Q2.1CO.landscape, echo=FALSE}
#Male.poly.Mouse.Table_BivData_4MM
#Curated_BivData_male

#glm1.SC.male.poly <- glm(mean_SC ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
#summary(glm1.SC.male.poly) 
#glm2.SC.male.poly <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
#summary(glm2.SC.male.poly)

#M1
glm.M1_Q2.1COnrm <- glm(norm.first.foci.pos ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
summary(glm.M1_Q2.1COnrm)

#M2
glm.M2_Q2.1COnrm <- glm(norm.first.foci.pos ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)

summary(glm.M2_Q2.1COnrm)

#Dom
glmQ2_M2.1COnrm_Dom <- glm(norm.first.foci.pos ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
#summary(glmQ2_M2.1COnrm_Dom)

#Musc
glmQ2_M2.1COnrm_Musc <- glm(norm.first.foci.pos ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
#summary(glmQ2_M2.1COnrm_Musc)

#Mol
glmQ2_M2.1COnrm_Mol <- glm(norm.first.foci.pos ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
#summary(glmQ2_M2.1COnrm_Mol)

#m3 (log)
Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
                                    ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
                                            ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)))

Log.glmQ2_M2.1COnrm <- glm(Rec.group ~ norm.first.foci.pos,
              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
#summary(Log.glmQ2_M2.1COnrm)

#Musc
Log.glmQ2_M2.1COnrm_Musc <- glm(Rec.group ~ norm.first.foci.pos,
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",], 
              family=binomial(link="logit"))
#summary(Log.glmQ2_M2.1COnrm_Musc)

#Mol
Log.glmQ2_M2.1COnrm_Mol <- glm(Rec.group ~ norm.first.foci.pos,
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],
              family=binomial(link="logit"))
#summary(Log.glmQ2_M2.1COnrm_Mol)

#biv level logistic regression
#Curated_BivData_male
#male.bivdata.1CO

male.bivdata.1CO$Rec.group <- ifelse(grepl("PWD", male.bivdata.1CO$strain), 1, 
                                    ifelse(grepl("MSM", male.bivdata.1CO$strain), 1, 
                                            ifelse(grepl("SKIVE", male.bivdata.1CO$strain), 1, 0)))

Log.Q2.1COnrm.F1.PER <- glm(Rec.group ~ PER_Foci_1,
              data=male.bivdata.1CO,  family=binomial(link="logit"))
summary(Log.glmQ2_M2.1COnrm)

#Musc
Log.Q2.F1_PER_Musc <- glm(Rec.group ~ PER_Foci_1,
              data=male.bivdata.1CO[male.bivdata.1CO$subsp == "Musc",], 
              family=binomial(link="logit"))
summary(Log.Q2.F1_PER_Musc)

#Mol
Log.Q2.F1_PER_Mol <- glm(Rec.group ~ PER_Foci_1,
              data=male.bivdata.1CO[male.bivdata.1CO$subsp == "Mol",],
              family=binomial(link="logit"))

summary(Log.glmQ2_M2.1COnrm_Mol)

The lm models

#### Summary

Not really a clear summary for the normalized 1CO position and rec groups.

@@
